---
title: "AFP Surveillance Monitoring — Demo"
date: today
date-format: "MMMM D, YYYY"
format:
  html:
    embed-resources: true
    page-layout: full
    theme: default
execute:
  echo: false
  message: false
  warning: false
---

```{r setup}
library(sirfunctions)
library(DT)
library(tidyverse)

end_date <- lubridate::ceiling_date(Sys.Date() %m-% months(1), unit = "month") %m-% days(1)

# Source repo functions ----
lapply(list.files("R"), \(fx) source(file.path("R", fx)))

```

```{r load-data}
raw_data <- get_all_polio_data(attach.spatial.data = FALSE)
afp_data <- raw_data$afp
pos <- raw_data$pos
```

```{r build-indicators}
# Build indicators — keep full list for metadata access
afp_cases_reported    <- build_afp_cases_reported(afp_data, end_date)
prop_60               <- build_prop_60_day_follow_up(afp_data)
afp_wpv_vdpv <- build_wpv_vdpv_timeliness_indicator(pos, end_date)

# Extract data and add risk category
afp_cases_reported_data <- afp_cases_reported$data |>
  sirfunctions:::add_risk_category(ctry_col = "place.admin.0") |>
  dplyr::select(-Region) |>
  dplyr::relocate(dplyr::any_of("SG Priority Level"), .after = "place.admin.0")

prop_60_data <- prop_60$data |>
  sirfunctions:::add_risk_category(ctry_col = "ctry") |>
  dplyr::select(-Region) |>
  dplyr::relocate(dplyr::any_of("SG Priority Level"), .after = "ctry")

afp_wpv_vdpv_data <- afp_wpv_vdpv$data |>
  sirfunctions:::add_risk_category(ctry_col = "ctry") |>
  dplyr::select(-Region) |>
  dplyr::relocate(dplyr::any_of("SG Priority Level"), .after = "ctry")
```


::: panel-tabset



## AFP Cases Reported

**Period:** `r afp_cases_reported$metadata$current_period_label`  
**Compared to:** `r afp_cases_reported$metadata$prior_period_label`  
**Threshold:** `r afp_cases_reported$metadata$threshold_rule`

```{r afp-cases-table}
datatable(
  afp_cases_reported_data |>
    dplyr::arrange(place.admin.0, month_label) |>
    dplyr::rename(
      Country      = "place.admin.0",
      Region       = "whoregion",
      `Month/Year` = "month_label",
      Cases        = "current_period_counts",
      `3yr Median` = "prior_median",
      `Yrs w/ Data`= "prior_yrs_w_data",
      `Lower 50%`  = "lower_50pct",
      `Upper 50%`  = "upper_50pct",
      Flag         = "flag"
    ),
  filter   = "top",
  extensions = "Buttons",
  options  = list(
    dom        = "Bfrtip",
    buttons    = c("copy", "csv", "excel"),
    pageLength = 15,
    scrollX    = TRUE,
    columnDefs = list(
      list(className = "dt-center", targets = "_all")
    )
  )
)
```

## Proportion 60-Day Follow-Up Done

**Recent period:** `r prop_60$metadata$recent_period_label`  
**Compared to:** `r prop_60$metadata$recent_prior_period_label`  
**Earlier period:** `r prop_60$metadata$earlier_period_label`  
**Compared to:** `r prop_60$metadata$earlier_prior_period_label`  
**Threshold:** `r prop_60$metadata$threshold_rule`

> `r prop_60$metadata$eligibility_note`

```{r prop-60-table}

datatable(
  prop_60_data |>
    dplyr::select(-window) |> 
    dplyr::arrange(ctry, current_period) |>
    dplyr::rename(
      Country              = "ctry",
      Region               = "whoregion",
      `Current Period`     = "current_period",
      `Prior Period`       = "prior_period",
      `Current Cases`      = "current_inad_cases",
      `Current Got 60d`    = "current_got_60day",
      `Current Prop (%)`   = "current_prop_60day",
      `Prior Cases`        = "prior_inad_cases",
      `Prior Got 60d`      = "prior_got_60day",
      `Prior Prop (%)`     = "prior_prop_60day",
      `% Change`           = "perc_change",
      Flag                 = "flag"
    ),
  filter     = "top",
  extensions = "Buttons",
  options    = list(
    dom        = "Bfrtip",
    buttons    = c("copy", "csv", "excel"),
    pageLength = 15,
    scrollX    = TRUE,
    columnDefs = list(
      list(className = "dt-center", targets = "_all")
    )
  )
)
```


## Timely AFP WPV/VDPV Detection

**Recent period:** `r afp_wpv_vdpv$metadata$recent_period_label`  
**Compared to:** `r afp_wpv_vdpv$metadata$recent_prior_period_label`  
**Earlier period:** `r afp_wpv_vdpv$metadata$earlier_period_label`  
**Compared to:** `r afp_wpv_vdpv$metadata$earlier_prior_period_label`  
**Threshold:** `r afp_wpv_vdpv$metadata$threshold_rule`

```{r wpv-vdpv-table}
datatable(
  afp_wpv_vdpv_data |>
    dplyr::select(-window) |> 
    dplyr::arrange(ctry, current_period) |>
    dplyr::rename(
      Country             = "ctry",
      Region              = "whoregion",
      `Current Period`    = "current_period",
      `Prior Period`      = "prior_period",
      `Current N`         = "current_count",
      `Current Median (days)` = "current_median_days",
      `Prior N`           = "prior_count",
      `Prior Median (days)`   = "prior_median_days",
      `% Change`          = "perc_change",
      Flag                = "flag"
    ),
  filter     = "top",
  extensions = "Buttons",
  options    = list(
    dom        = "Bfrtip",
    buttons    = c("copy", "csv", "excel"),
    pageLength = 15,
    scrollX    = TRUE,
    columnDefs = list(
      list(className = "dt-center", targets = "_all")
    )
  )
)
```

:::
