---
title: "Surveillance Group Monitoring Report"
author: Mervin Keith Cuadera
date: today
date-format: "MMMM D, YYYY"  # e.g., September 24, 2025
format: 
  html:
    page-layout: custom
    embed-resources: true
output-file: 
---

```{r}
#| echo: false
#| include: false

library(sirfunctions)
library(DT)
library(tidyverse)

use_cache <- TRUE # Change if creating report using previous run or new run
end_date <- "2025-12-31" # Change as needed. Only works for new runs

if (use_cache) {
  if (!dir.exists("data_cache")) {
    dir.create("data_cache")
  }
  
  if (length(list.files("data_cache")) == 0) {
    cli::cli_alert_warning("No data found in cache. Performing a new run.")
    source("generate_data_for_report.R")
  } else {
    load("data_cache/cache.rda")
  }
} else {
  source("generate_data_for_report.R")
}

tile_plots <- list.files("R/", "_tile_plot.R")
lapply(tile_plots, \(x) source(file.path("R", x)))


```

## System-Level Performance

### AFP Indicators
::: panel-tabset

#### AFP Indicators
```{r, fig.width=14, fig.height=8}
#| echo: false
#| message: false
#| warning: false

generate_afp_tile_plot(afp_cases_reported, prop_60, lab_pending, prop_classified,
                       afp_wpv_vdpv, negative_lab_processing, afp_shipment_timeliness,
                       afp_lab_processing, end_date = Sys.Date(), lab_end_date = max_lab_date, who_region = "AFRO")

generate_es_tile_plot(es_shipment, es_wpv_vdpv, es_sites, es_site_samples, end_date = Sys.Date(), who_region = "AFRO")

```
:::

## AFP Surveillance Indicators

::: panel-tabset

### Number of AFP cases reported

```{r}
#| echo: false

datatable(
  afp_cases_reported |>
    dplyr::rename(
      Region = "whoregion",
      Country = "place.admin.0",
      Month = "month"
    ),
  filter = "top",
  extensions = 'Buttons',
  options = list(
    dom = 'Bfrtip',
    buttons = c('copy', 'csv', 'excel'),
    pageLength = 10
  )
)
```

### Proportion of 60 days follow-up done
```{r}
#| echo: false

datatable(
  prop_60 |>
    dplyr::rename_with(dplyr::recode,
                ctry = "Country",
                quarter = "Quarter",
                semester = "Semester",
                month = "Month",
                comparison = "Difference",
                trend = "Trend"),
  filter = "top",
  extensions = 'Buttons',
  options = list(
    dom = 'Bfrtip',
    buttons = c('copy', 'csv', 'excel'),
    pageLength = 10
  )
)
```

### Proportion of lab pending

Shows the proportion of lab samples not yet received in the culturing lab, with pending classification.
```{r}
#| echo: false

datatable(
  lab_pending |>
    dplyr::rename(
      Country = "country",
      Region = "whoregion",
      Year = "year",
      Month = "month",
      `Pending Samples` = "pending_samples",
      `% Lab Samples` = "prop_label"
    ),
  filter = "top",
  extensions = 'Buttons',
  options = list(
    dom = 'Bfrtip',
    buttons = c('copy', 'csv', 'excel'),
    pageLength = 10
  )
)

```

### Proportion of cases classified
Proportion of cases with inadequate samples classified by the ERC.
```{r}
#| echo: false

datatable(
  prop_classified |>
    dplyr::rename(
      Country = "ctry",
      Quarter = "quarter",
      `% Difference` = "diff",
      Trend = "trend",
      `Current Year Performance` = "performance"
    ),
  filter = "top",
  extensions = 'Buttons',
  options = list(
    dom = 'Bfrtip',
    buttons = c('copy', 'csv', 'excel'),
    pageLength = 10
  )
)
```

### Timeliness of detection for WPV/VDPV samples
```{r}
#| echo: false

datatable(
  afp_wpv_vdpv |>
    dplyr::rename_with(recode,
                country = "Country",
                whoregion = "Region",
                month = "Month",
                quarter = "Quarter",
                comparison = "Difference",
                trend = "Trend"),
  filter = "top",
  extensions = 'Buttons',
  options = list(
    dom = 'Bfrtip',
    buttons = c('copy', 'csv', 'excel'),
    pageLength = 10
  )
)

```

### Timeliness of detection for negative samples
From date of collection to final cell culture results.
```{r}
#| echo: false

datatable(
  negative_lab_processing |>
    dplyr::rename(
      Country = "country",
      Region = "whoregion",
      Category = "culture.itd.cat",
      Month = "month",
      Difference = "comparison",
      Trend = "trend"
    ),
  filter = "top",
  extensions = 'Buttons',
  options = list(
    dom = 'Bfrtip',
    buttons = c('copy', 'csv', 'excel'),
    pageLength = 10
  )
)

```

### Timeliness of stool specimen shipment
```{r}
#| echo: false

datatable(
  afp_shipment_timeliness |>
    dplyr::rename_with(recode,
      country = "Country",
      whoregion = "Region",
      month = "Month",
      comparison = "Difference",
      culture.itd.cat = "Shipment Type",
      trend = "Trend"
    ),
  filter = "top",
  extensions = 'Buttons',
  options = list(
    dom = 'Bfrtip',
    buttons = c('copy', 'csv', 'excel'),
    pageLength = 10
  )
)

```

### Timeliness of lab processing of AFP samples
```{r}
#| echo: false

datatable(
  afp_lab_processing |>
    dplyr::rename_with(recode,
      country = "Country",
      month = "Month",
      quarter = "Quarter",
      whoregion = "Region",
      comparison = "Difference",
      trend = "Trend"
    ),
  filter = "top",
  extensions = 'Buttons',
  options = list(
    dom = 'Bfrtip',
    buttons = c('copy', 'csv', 'excel'),
    pageLength = 10
  )
)
```

### Proportion of inadequates with contact sampling
```{r}
#| echo: false

datatable(
  prop_inad |>
    dplyr::rename(
      Country = "ctry",
      Quarter = "quarter",
      Region = "whoregion",
      `Difference, %` = "comparison",
      Trend = "trend"
    ),
  filter = "top",
  extensions = 'Buttons',
  options = list(
    dom = 'Bfrtip',
    buttons = c('copy', 'csv', 'excel'),
    pageLength = 10
  )
)

```
:::

## Environmental Surveillance Indicators

::: panel-tabset
### Timeliness of ES sample shipment

```{r}
#| label: es-lab-shipment
#| echo: false
es_shipment <- es_shipment |>
  dplyr::rename(
    Month = "month",
    Country = "country",
    `Lab Type` = "es.lab.type",
    Region = "who.region",
    Difference = "diff",
    Trend = "trend",
    `Current Year Timeliness` = "current_year_timeliness",
    `Trend Summary` = "trend_summary"
  )

datatable(
  es_shipment,
  filter = "top",
  extensions = 'Buttons',
  options = list(
    dom = 'Bfrtip',
    buttons = c('copy', 'csv', 'excel'),
    pageLength = 10
  )
)

```

### Timeliness of detection for WPV/VDPV-ES

```{r}
#| label: es-wpv-vdpv
#| echo: false

es_wpv_vdpv <- es_wpv_vdpv |>
  dplyr::rename(
    Month = "month",
    Country = "country",
    `Lab Type` = "es.lab.type",
    Region = "who.region",
    Difference = "diff",
    Trend = "trend",
    `Current Year Timeliness` = "current_year_timeliness",
    `Trend Summary` = "trend_summary"
  )

datatable(
  es_wpv_vdpv,
  filter = "top",
  extensions = 'Buttons',
  options = list(
    dom = 'Bfrtip',
    buttons = c('copy', 'csv', 'excel'),
    pageLength = 10
  )
)
```

### Number of operational ES sites per country

```{r}
#| label: es-sites
#| echo: false

es_sites <- es_sites |>
  dplyr::rename(
    Country = "ctry",
    Difference = "comparison",
    `Difference, %` = "prop_diff",
    Trend = "trend"
  )

datatable(
  es_sites,
  filter = "top",
  extensions = 'Buttons',
  options = list(
    dom = 'Bfrtip',
    buttons = c('copy', 'csv', 'excel'),
    pageLength = 10
  )
)

```

### Number of samples per sites

```{r}
#| label: es-site-samples
#| include: true
#| echo: false

es_site_samples <- es_site_samples |>
  dplyr::rename(
    Country = "ADM0_NAME",
    `Site Name` = "site.name",
    `Site Age` = "site_age",
    `Samples In Last 12 Months` = "n_samples_12_mo",
    `POLIS Site Status` = "site.status",
    `Earliest Collection` = "earliest_collection",
    `Last Collection` = "last_collection",
    `Days Since Last Collection` = "days_since_last_collection",
    `Collection In Two Months` = "collection_two_mo",
    `Last Detection` = "last_detection",
    `Days Since Last Detection` = "days_since_last_det",
    `Detection In Two Months` = "detection_two_mo",
    `Is Active Site` = "active_site"
  )

datatable(
  es_site_samples,
  filter = "top",
  extensions = 'Buttons',
  options = list(
    dom = 'Bfrtip',
    buttons = c('copy', 'csv', 'excel'),
    pageLength = 10
  )
)

```
:::

## Lab Surveillance Indicators
`{r} max_date_notif`

::: panel-tabset
### Timeliness of virus isolation results

```{r}
#| include: false
#| echo: false

culture_lab_intervals <- culture_lab_intervals |>
  dplyr::rename_with(recode,
                     culture.itd.lab = "Culture/ITD Lab",
                     comparison = "Difference",
                     prop_diff = "Difference, %",
                     trend = "Trend")

interval_summary <- culture_lab_intervals |> 
    dplyr::filter(interval == "days.lab.culture") |>
    dplyr::select(-interval)
top_10 <- interval_summary |> 
  filter(Trend == "Increase") |> 
  arrange(desc(Difference)) |>
  select(-Trend) |>
  slice_head(n = 10)
bottom_10 <-interval_summary |> 
  filter(Trend == "Decrease") |> 
  arrange(Difference) |>
  select(-Trend) |>
  slice_head(n = 10)

summary_string_top <- glue::glue_data(
  top_10,
  "{str_to_title(`Culture/ITD Lab`)} ({round(Difference, 1)})"
)
summary_string_bottom <- glue::glue_data(
  bottom_10,
  "{str_to_title(`Culture/ITD Lab`)} ({round(Difference, 1)})"
)

# Combine into a single sentence
final_summary_top <- paste0("The top labs with increased median timeliness compared to the three year median (difference in days shown) are: ", paste(summary_string_top, collapse = ", "), ".")
final_summary_bottom <- paste0("The top 10 labs with the biggest improvement in median timeliness compared to the three year median (difference in days shown) are: ", paste(summary_string_bottom, collapse = ", "), ".")
comparible <- paste("There were", interval_summary |> 
                      filter(Trend == "Same") |> 
                      nrow(), "labss the stayed in the same levels and", 
                    interval_summary |> 
                      filter(Trend == "No data available for both years") |> 
                      nrow(), "labs with missing 3-year medians or 2025 medians.")
overall <- paste0("Overall, ", 
                  interval_summary |> 
                    filter(!is.na(`Culture/ITD Lab`), 
                           Trend != "No data available for both years") |>
                    pull(`Culture/ITD Lab`) |> 
                    unique() |> 
                    length(), " labs had data in the lab dataset. ",
                  interval_summary |>
                    filter(!is.na(`Culture/ITD Lab`)) |>
                    select(4) |> 
                    filter(if_all(1, \(x) !is.na(x))) |>
                    nrow(), " labs had medians for the current year and ",
                  interval_summary |> 
                    filter(!is.na(`Culture/ITD Lab`)) |>
                    select(3) |> 
                    filter(if_all(1, \(x) !is.na(x))) |>
                    nrow(), " labs had 3-year medians."
                  )

```

#### Summary

`r final_summary_top`

`r final_summary_bottom`

`r comparible`

`r overall`

```{r}
#| include: true
#| echo: false

datatable(interval_summary,
  filter = "top",
  extensions = 'Buttons',
  options = list(
    dom = 'Bfrtip',
    buttons = c('copy', 'csv', 'excel'),
    pageLength = 10
  )
)
```

### Timeliness of ITD results

```{r}
#| include: false
#| echo: false

interval_summary <- culture_lab_intervals |> 
    dplyr::filter(interval == "days.culture.itd") |>
    dplyr::select(-interval)
top_10 <- interval_summary |> 
  filter(Trend == "Increase") |> 
  arrange(desc(Difference)) |>
  select(-Trend) |>
  slice_head(n = 10)
bottom_10 <-interval_summary |> 
  filter(Trend == "Decrease") |> 
  arrange(Difference) |>
  select(-Trend) |>
  slice_head(n = 10)

summary_string_top <- glue::glue_data(
  top_10,
  "{str_to_title(`Culture/ITD Lab`)} ({round(Difference, 1)})"
)
summary_string_bottom <- glue::glue_data(
  bottom_10,
  "{str_to_title(`Culture/ITD Lab`)} ({round(Difference, 1)})"
)

# Combine into a single sentence
final_summary_top <- paste0("The top labs with increased median timeliness compared to the three year median (difference in days shown) are: ", paste(summary_string_top, collapse = ", "), ".")
final_summary_bottom <- paste0("The top 10 labs with the biggest improvement in median timeliness compared to the three year median (difference in days shown) are: ", paste(summary_string_bottom, collapse = ", "), ".")
comparible <- paste("There were", interval_summary |> 
                      filter(Trend == "Same") |> 
                      nrow(), "labss the stayed in the same levels and", 
                    interval_summary |> 
                      filter(Trend == "No data available for both years") |> 
                      nrow(), "labs with missing 3-year medians or 2025 medians.")
overall <- paste0("Overall, ", 
                  interval_summary |> 
                    filter(!is.na(`Culture/ITD Lab`), 
                           Trend != "No data available for both years") |>
                    pull(`Culture/ITD Lab`) |> 
                    unique() |> 
                    length(), " labs had data in the lab dataset. ",
                  interval_summary |>
                    filter(!is.na(`Culture/ITD Lab`)) |>
                    select(4) |> 
                    filter(if_all(1, \(x) !is.na(x))) |>
                    nrow(), " labs had medians for the current year and ",
                  interval_summary |> 
                    filter(!is.na(`Culture/ITD Lab`)) |>
                    select(3) |> 
                    filter(if_all(1, \(x) !is.na(x))) |>
                    nrow(), " labs had 3-year medians."
                  )
```

#### Summary

`r final_summary_top`

`r final_summary_bottom`

`r comparible`

`r overall`

```{r}
#| include: true
#| echo: false

datatable(
  culture_lab_intervals |> 
    dplyr::filter(interval == "days.culture.itd") |>
    dplyr::select(-interval),
  filter = "top",
  extensions = 'Buttons',
  options = list(
    dom = 'Bfrtip',
    buttons = c('copy', 'csv', 'excel'),
    pageLength = 10
  )
)
```

### Timeliness of shipment for sequencing

```{r}
#| include: false
#| echo: false

interval_summary <- culture_lab_intervals |> 
    dplyr::filter(interval == "days.seq.ship") |>
    dplyr::select(-interval)
top_10 <- interval_summary |> 
  filter(Trend == "Increase") |> 
  arrange(desc(Difference)) |>
  select(-Trend) |>
  slice_head(n = 10)
bottom_10 <-interval_summary |> 
  filter(Trend == "Decrease") |> 
  arrange(Difference) |>
  select(-Trend) |>
  slice_head(n = 10)

summary_string_top <- glue::glue_data(
  top_10,
  "{str_to_title(`Culture/ITD Lab`)} ({round(Difference, 1)})"
)
summary_string_bottom <- glue::glue_data(
  bottom_10,
  "{str_to_title(`Culture/ITD Lab`)} ({round(Difference, 1)})"
)

# Combine into a single sentence
final_summary_top <- paste0("The top labs with increased median timeliness compared to the three year median (difference in days shown) are: ", paste(summary_string_top, collapse = ", "), ".")
final_summary_bottom <- paste0("The top 10 labs with the biggest improvement in median timeliness compared to the three year median (difference in days shown) are: ", paste(summary_string_bottom, collapse = ", "), ".")
comparible <- paste("There were", interval_summary |> 
                      filter(Trend == "Same") |> 
                      nrow(), "labss the stayed in the same levels and", 
                    interval_summary |> 
                      filter(Trend == "No data available for both years") |> 
                      nrow(), "labs with missing 3-year medians or 2025 medians.")
overall <- paste0("Overall, ", 
                  interval_summary |> 
                    filter(!is.na(`Culture/ITD Lab`), 
                           Trend != "No data available for both years") |>
                    pull(`Culture/ITD Lab`) |> 
                    unique() |> 
                    length(), " labs had data in the lab dataset. ",
                  interval_summary |>
                    filter(!is.na(`Culture/ITD Lab`)) |>
                    select(4) |> 
                    filter(if_all(1, \(x) !is.na(x))) |>
                    nrow(), " labs had medians for the current year and ",
                  interval_summary |> 
                    filter(!is.na(`Culture/ITD Lab`)) |>
                    select(3) |> 
                    filter(if_all(1, \(x) !is.na(x))) |>
                    nrow(), " labs had 3-year medians."
                  )

```

#### Summary

`r final_summary_top`

`r final_summary_bottom`

`r comparible`

`r overall`

```{r}
#| include: true
#| echo: false

datatable(
  culture_lab_intervals |> 
    dplyr::filter(interval == "days.seq.ship") |>
    dplyr::select(-interval),
  filter = "top",
  extensions = 'Buttons',
  options = list(
    dom = 'Bfrtip',
    buttons = c('copy', 'csv', 'excel'),
    pageLength = 10
  )
)
```

### Timeliness of sequencing results

```{r}
#| include: false
#| echo: false
seq_lab_interval <- seq_lab_interval |>
  dplyr::rename_with(recode,
                     seq.lab = "Sequencing Lab",
                     comparison = "Difference",
                     prop_diff = "Difference, %",
                     trend = "Trend")

interval_summary <- seq_lab_interval |> 
    dplyr::filter(interval == "days.seq.rec.res") |>
    dplyr::select(-interval)
top_10 <- interval_summary |> 
  filter(Trend == "Increase") |> 
  arrange(desc(Difference)) |>
  select(-Trend) |>
  slice_head(n = 10)
bottom_10 <-interval_summary |> 
  filter(Trend == "Decrease") |> 
  arrange(Difference) |>
  select(-Trend) |>
  slice_head(n = 10)

summary_string_top <- glue::glue_data(
  top_10,
  "{str_to_title(`Sequencing Lab`)} ({round(Difference, 1)})"
)
summary_string_bottom <- glue::glue_data(
  bottom_10,
  "{str_to_title(`Sequencing Lab`)} ({round(Difference, 1)})"
)

# Combine into a single sentence
final_summary_top <- paste0("The top labs with increased median timeliness compared to the three year median (difference in days shown) are: ", paste(summary_string_top, collapse = ", "), ".")
final_summary_bottom <- paste0("The top 10 labs with the biggest improvement in median timeliness compared to the three year median (difference in days shown) are: ", paste(summary_string_bottom, collapse = ", "), ".")
comparible <- paste("There were", interval_summary |> 
                      filter(Trend == "Same") |> 
                      nrow(), "labss the stayed in the same levels and", 
                    interval_summary |> 
                      filter(Trend == "No data available for both years") |> 
                      nrow(), "labs with missing 3-year medians or 2025 medians.")
overall <- paste0("Overall, ", 
                  interval_summary |> 
                    filter(!is.na(`Sequencing Lab`), 
                           Trend != "No data available for both years") |>
                    pull(`Sequencing Lab`) |> 
                    unique() |> 
                    length(), " labs had data in the lab dataset. ",
                  interval_summary |>
                    filter(!is.na(`Sequencing Lab`)) |>
                    select(4) |> 
                    filter(if_all(1, \(x) !is.na(x))) |>
                    nrow(), " labs had medians for the current year and ",
                  interval_summary |> 
                    filter(!is.na(`Sequencing Lab`)) |>
                    select(3) |> 
                    filter(if_all(1, \(x) !is.na(x))) |>
                    nrow(), " labs had 3-year medians."
                  )

```

#### Summary

`r final_summary_top`

`r final_summary_bottom`

`r comparible`

`r overall`

```{r}
#| include: true
#| echo: false

datatable(
  seq_lab_interval |> 
    dplyr::filter(interval == "days.seq.rec.res") |>
    dplyr::select(-interval),
  filter = "top",
  extensions = 'Buttons',
  options = list(
    dom = 'Bfrtip',
    buttons = c('copy', 'csv', 'excel'),
    pageLength = 10
  )
)
```

### Laboratory workload

```{r}
#| include: false
#| echo: false

lab_workload <- lab_workload |>
  dplyr::rename(
    `Culture/ITD Lab` = "culture.itd.lab",
    Month = "month",
    `Countries Served` = "from_ctry",
    Difference = "comparison",
    `Difference (%)` = "comparison_pct",
    Trend = "trend"
  )

last_month <- lab_workload |>
  filter(Month == lubridate::month(lubridate::month(max_lab_date) - 1, TRUE))

top_10 <- last_month |>
  filter(Trend == "Increase") |>
  arrange(desc(Difference)) |>
  slice_head(n = 10)

bottom_10 <- last_month |> 
  filter(Trend == "Decrease") |> 
  arrange(Difference) |>
  select(-Trend) |>
  slice_head(n = 10)

summary_string_top <- glue::glue_data(
  top_10,
  "{str_to_title(`Culture/ITD Lab`)} ({round(Difference, 1)})"
)
summary_string_bottom <- glue::glue_data(
  bottom_10,
  "{str_to_title(`Culture/ITD Lab`)} ({round(Difference, 1)})"
)

# Combine into a single sentence
final_summary_top <- paste0("The top 10 labs with increased lab samples compared to the previous year (difference in samples shown) are: ", paste(summary_string_top, collapse = ", "), ".")
final_summary_bottom <- paste0("The top 10 labs with reduced lab samples compared to the previous year (difference in samples shown) are: ", paste(summary_string_bottom, collapse = ", "), ".")
comparible <- paste("There were", last_month |> 
                      filter(Trend == "Same") |> nrow(), "labs the stayed in the same levels and", last_month |> 
                      filter(Trend == "No data available for both years") |> 
                      nrow(), "labs with missing data for the current year and previous year.")
overall <- paste0("Overall, ", 
                  last_month |> 
                    filter(if_any(5, \(x) !is.na(x))) |> 
                    nrow(), 
                  " labs had data from the previous month of the current year using the latest month available in the lab dataset.")

```

#### Summary

**For the previous month** 

`r final_summary_top` 

`r final_summary_bottom` 

`r comparible`

`r overall`

```{r}
#| include: true
#| echo: false
datatable(
  lab_workload,
  filter = "top",
  extensions = 'Buttons',
  options = list(
    dom = 'Bfrtip',
    buttons = c('copy', 'csv', 'excel'),
    pageLength = 10
  )
)
```
:::
