---
title: "GPEI Surveillance Group Monitoring Report"
date: today
date-format: "MMMM D, YYYY"
format:
  html:
    embed-resources: true
    page-layout: full
    theme: default
execute:
  echo: false
  message: false
  warning: false
---


```{r setup}
#| echo: false
#| include: false
#| warning: false

library(sirfunctions)
library(DT)
library(tidyverse)
library(plotly)
library(crosstalk)
library(cowplot)

# Set end date for report
end_date <- lubridate::ceiling_date(Sys.Date() %m-% months(2), unit = "month") %m-% days(1)

# Get report data from existing tables or by sourcing generate_data_for_report.R
use_existing_data <- FALSE  # set TRUE to skip data reload during QMD development

if (use_existing_data) {
  load("datatables/datatables.rda")
} else {
  source("generate_data_for_report.R")
}
# Note: build_ and create_ functions are sourced inside generate_data_for_report.R

#function for making month column sortable chronologically instead of alphabetically (should eventually be moved to a helper function )
make_sortable_month_col <- function(data, 
                                       month_col_name = "Month",
                                       filter = "top",
                                       buttons = c("copy", "csv", "excel"),
                                       pageLength = 15,
                                       scrollX = TRUE) {
  
  month_col <- which(names(data) == month_col_name) - 1
  
  if (length(month_col) == 0) {
    warning(paste0("Column '", month_col_name, "' not found — sorting fix not applied."))
  }
  
  datatable(
    data,
    rownames = FALSE,
    filter = filter,
    extensions = "Buttons",
    options = list(
      dom = "Bfrtip",
      buttons = buttons,
      pageLength = pageLength,
      scrollX = scrollX,
      columnDefs = list(
        list(
          targets = month_col,
          type = "num",
          render = DT::JS(
            "function(data, type, row) {",
            "  if (type === 'sort' || type === 'type') {",
            "    var d = new Date('01 ' + data);",
            "    return d.getTime();",
            "  }",
            "  return data;",
            "}"
          )
        )
      )
    )
  )
}

```

::: panel-tabset

## Overview

::: panel-tabset

### Summary

End date: **`r format(end_date, "%B %d, %Y")`**

The objective of the monthly SG Monitoring Report is as an early warning for potential disruption in the surveillance system.  
Monitoring Flags are presented at the country level across AFP, ES, and Lab. 


:::

## AFP and ES Summary Tables & Visuals

::: panel-tabset

### Summary Table - Region

The total number of countries in the denominator may change between quarters. See individual flag tabs for definition of incomplete/insufficient data.


```{r region-summary-table-alt}
#| echo: false

region_table_alt <- region_table |>
    dplyr::select(whoregion, flagname, string_Earlier, string_Current, countries_below_Current, countries_incomplete_Current) |>
    dplyr::arrange(whoregion, flagname) |>
    dplyr::rename(
      Region = whoregion,
      Flag=flagname,
      `% of Countries Flagged - Earlier` = string_Earlier,
      `% of Countries Flagged` = string_Current,
      `Countries Flagged` = countries_below_Current,
      `Countries with insufficient data` = countries_incomplete_Current,
    ) |>
  #need to create a unique row for dynamic filtering
  dplyr::mutate(row_key = paste0(Region, "_", Flag)) |>
  SharedData$new(key = ~row_key)


filter_select("region-table", "Region: ", 
              region_table_alt,
              ~Region,
              multiple = FALSE,
              allLevels = FALSE)

datatable(
  region_table_alt, 
  filter = "top",
  extensions = 'Buttons',
  rownames=FALSE,
  options = list(
    dom = 'Bfrtip',
    buttons = c('copy', 'csv', 'excel'),
    pageLength = 12,
    #need to remove the row_key column we created for dynamic filtering
    columnDefs = list(
      list(visible = FALSE, targets = which(names(region_table_alt$data()) == "row_key") - 1)
    )
  )
)

```

### Flagged Countries, Monthly Flags

```{r country-summary-table-m}
#| echo: false

country_table_m<-country_table_monthly     |>
    dplyr::select(Flag, whoregion,Country,sg_priority_level,dplyr::matches("[0-9]"),months_below) |>
    dplyr::rename(
      Region = whoregion,
      `SG Priority Level`=sg_priority_level,
      `Months Below Target` = months_below,
    )|>
  #need to create a unique row for dynamic filtering
  dplyr::mutate(row_key = paste0(Country, "_", Flag)) |>
  SharedData$new(key = ~row_key)

filter_select("country-table-m-region", "Region: ", 
              country_table_m,
              ~Region,
              multiple = FALSE,
              allLevels = FALSE)

filter_select("country-table-m-flag", "Flag: ", 
              country_table_m,
              ~Flag,
              multiple = FALSE,
              allLevels = FALSE)

datatable(
  country_table_m, 
  filter = "top",
  extensions = 'Buttons',
  rownames=FALSE,
  options = list(
    dom = 'Bfrtip',
    buttons = c('copy', 'csv', 'excel'),
    pageLength = 20,
    #need to remove the row_key column we created for dynamic filtering
    columnDefs = list(
      list(visible = FALSE, targets = which(names(country_table_m$data()) == "row_key") - 1)
    )
  )
)


```

### Flagged Countries, Quarterly Flags

```{r country-summary-table-q}
#| echo: false

country_table_q<-country_table_quarterly    |>
    dplyr::select(Flag,whoregion,Country,sg_priority_level,Months) |>
    dplyr::rename(
      Region = whoregion,
      `SG Priority Level`=sg_priority_level    )|>
  #need to create a unique row for dynamic filtering
  dplyr::mutate(row_key = paste0(Country, "_", Flag)) |>
  SharedData$new(key = ~row_key)

filter_select("country-table-q-region", "Region: ", 
              country_table_q,
              ~Region,
              multiple = FALSE,
              allLevels = FALSE)

filter_select("country-table-q-flag", "Flag: ", 
              country_table_q,
              ~Flag,
              multiple = FALSE,
              allLevels = FALSE)

datatable(
  country_table_q, 
  filter = "top",
  extensions = 'Buttons',
  rownames=FALSE,
  options = list(
    dom = 'Bfrtip',
    buttons = c('copy', 'csv', 'excel'),
    pageLength = 20,
    #need to remove the row_key column we created for dynamic filtering
    columnDefs = list(
      list(visible = FALSE, targets = which(names(country_table_q$data()) == "row_key") - 1)
    )
  )
)


```

### Flagged Countries, Flags with other time windows

See individual flag tabs for specific time window.

```{r country-summary-table-n}
#| echo: false

country_table_n<-country_table_noperiod    |>
    dplyr::select(Flag, whoregion, Country, sg_priority_level, Pct) |>
    dplyr::rename(
      Region = whoregion,
      `SG Priority Level`=sg_priority_level,
      `Proportion`=Pct
    )|>
  #need to create a unique row for dynamic filtering
  dplyr::mutate(row_key = paste0(Country, "_", Flag)) |>
  SharedData$new(key = ~row_key)

filter_select("country-table-n-region", "Region: ", 
              country_table_n,
              ~Region,
              multiple = FALSE,
              allLevels = FALSE)

filter_select("country-table-n-flag", "Flag: ", 
              country_table_n,
              ~Flag,
              multiple = FALSE,
              allLevels = FALSE)

datatable(
  country_table_n, 
  filter = "top",
  extensions = 'Buttons',
  rownames=FALSE,
  options = list(
    dom = 'Bfrtip',
    buttons = c('copy', 'csv', 'excel'),
    pageLength = 20,
    #need to remove the row_key column we created for dynamic filtering
    columnDefs = list(
      list(visible = FALSE, targets = which(names(country_table_n$data()) == "row_key") - 1)
    )
  )
)


```

:::


## Lab Summary Tables & Visuals

::: panel-tabset

### Lab Summary Table - Region

The total number of labs in the denominator may change between quarters. See individual flag tabs for definition of incomplete/insufficient data.
Labs are assigned to WHO regions based on location.

```{r lab-region-summary-table}
#| echo: false

lab_region_table_wide <- lab_region_table |>
    dplyr::select(whoregion, flagname, string_Earlier, string_Current, countries_below_Current, countries_incomplete_Current) |>
    dplyr::arrange(whoregion, flagname) |>
    dplyr::rename(
      Region = whoregion,
      Flag=flagname,
      `% of Labs Flagged - Earlier` = string_Earlier,
      `% of Labs Flagged` = string_Current,
      `Labs Flagged` = countries_below_Current,
      `Labs with insufficient data` = countries_incomplete_Current,
    ) |>
  #need to create a unique row for dynamic filtering
  dplyr::mutate(row_key = paste0(Region, "_", Flag)) |>
  SharedData$new(key = ~row_key)


filter_select("lab-region-table", "Region: ", 
              lab_region_table_wide,
              ~Region,
              multiple = FALSE,
              allLevels = FALSE)

datatable(
  lab_region_table_wide, 
  filter = "top",
  extensions = 'Buttons',
  rownames=FALSE,
  options = list(
    dom = 'Bfrtip',
    buttons = c('copy', 'csv', 'excel'),
    pageLength = 12,
    #need to remove the row_key column we created for dynamic filtering
    columnDefs = list(
      list(visible = FALSE, targets = which(names(lab_region_table_wide$data()) == "row_key") - 1)
    )
  )
)

```


### Flagged Labs, Monthly Flags

```{r lab-country-summary-table-m}
#| echo: false

lab_country_table_monthly<-lab_country_table_monthly     |>
    dplyr::select(Flag, whoregion, Country, dplyr::matches("[0-9]"),months_below) |>
    dplyr::rename(
      Region = whoregion,
      Lab = Country,
      `Months Below Target` = months_below
    )|>
  #need to create a unique row for dynamic filtering
  dplyr::mutate(row_key = paste0(Lab, "_", Flag)) |>
  SharedData$new(key = ~row_key)

filter_select("lab-country-table-m-region", "Region: ", 
              lab_country_table_monthly,
              ~Region,
              multiple = FALSE,
              allLevels = FALSE)

filter_select("lab-country-table-m-flag", "Flag: ", 
              lab_country_table_monthly,
              ~Flag,
              multiple = FALSE,
              allLevels = FALSE)

datatable(
  lab_country_table_monthly, 
  filter = "top",
  extensions = 'Buttons',
  rownames=FALSE,
  options = list(
    dom = 'Bfrtip',
    buttons = c('copy', 'csv', 'excel'),
    pageLength = 20,
    #need to remove the row_key column we created for dynamic filtering
    columnDefs = list(
      list(visible = FALSE, targets = which(names(lab_country_table_monthly$data()) == "row_key") - 1)
    )
  )
)


```

### Flagged Labs, Quarterly Flags

```{r lab-country-summary-table-q}
#| echo: false

lab_country_table_quarterly<-lab_country_table_quarterly    |>
    dplyr::select(Flag, whoregion, Country, Months) |>
    dplyr::rename(
      Region = whoregion,
      Lab=Country)|>
  #need to create a unique row for dynamic filtering
  dplyr::mutate(row_key = paste0(Lab, "_", Flag)) |>
  SharedData$new(key = ~row_key)

filter_select("lab-country-table-q-region", "Region: ", 
              lab_country_table_quarterly,
              ~Region,
              multiple = FALSE,
              allLevels = FALSE)

filter_select("lab-country-table-q-flag", "Flag: ", 
              lab_country_table_quarterly,
              ~Flag,
              multiple = FALSE,
              allLevels = FALSE)

datatable(
  lab_country_table_quarterly, 
  filter = "top",
  extensions = 'Buttons',
  rownames=FALSE,
  options = list(
    dom = 'Bfrtip',
    buttons = c('copy', 'csv', 'excel'),
    pageLength = 20,
    #need to remove the row_key column we created for dynamic filtering
    columnDefs = list(
      list(visible = FALSE, targets = which(names(lab_country_table_quarterly$data()) == "row_key") - 1)
    )
  )
)


```

:::

## AFP Monitoring Flags

::: panel-tabset

### Number of AFP Cases Reported

**Period of focus:** `r afp_cases_reported$metadata$current_period_label`  
**Comparison periods:** `r afp_cases_reported$metadata$prior_period_label`  
**Threshold for above or below target:** `r afp_cases_reported$metadata$threshold_rule`  

```{r afp-cases-table}
afp_cases_reported$data |>
    dplyr::arrange(place.admin.0, lubridate::my(month_label)) |>
    dplyr::rename(
      Country = place.admin.0,
      Region = whoregion,
      `SG Priority Level` = sg_priority_level,
      Month = month_label,
      Cases = current_period_counts,
      `3yr Median` = prior_3yr_median,
      `Yrs w/ Data` = prior_yrs_w_data,
      `Percent Change` = perc_change,
      Flag = flag
    ) |>
  make_sortable_month_col()
```

### Proportion 60-Day Follow-Up Completed

**Period of focus:** `r afp_prop_60$metadata$recent_period_label`  
**Earlier period (for context):** `r afp_prop_60$metadata$earlier_period_label`  
**Eligibility:** `r afp_prop_60$metadata$eligibility_note`  
**Threshold:** `r afp_prop_60$metadata$threshold_rule`

```{r prop-60-table}
datatable(
  afp_prop_60$data |>
    dplyr::arrange(ctry, period) |>
    dplyr::rename(
      Country = ctry,
      Region = whoregion,
      `SG Priority Level` = sg_priority_level,
      Period = period,
      `Inadequate Cases` = inad_cases,
      `Got 60d` = got_60day,
      `% Got 60d` = prop_60day,
      Flag = flag
    ),
  rownames=FALSE,
  filter = "top", 
  extensions = "Buttons",
  options = list(
    dom = "Bfrtip", 
    buttons = c("copy", "csv", "excel"),
    pageLength = 15, 
    scrollX = TRUE
    )
)
```

### Proportion Inadequate Cases Unclassified

**Eligibility:** `r afp_prop_inad_classified$metadata$eligibility_note`  
**Threshold:** `r afp_prop_inad_classified$metadata$threshold_rule`

```{r prop-classified-table}
datatable(
  afp_prop_inad_classified$data |>
    dplyr::arrange(ctry) |>
    dplyr::rename(
      Country = ctry,
      Region = whoregion,
      `SG Priority Level` = sg_priority_level,
      `Inad Cases` = inad_cases,
      `# Classified` = n_classified,
      `# Unclassified` = n_unclassified,
      `% Unclassified` = prop_unclassified,
      Flag = flag
    ),
  rownames=FALSE,
  filter = "top", 
  extensions = "Buttons",
  options = list(
    dom = "Bfrtip", 
    buttons = c("copy", "csv", "excel"),
    pageLength = 15, 
    scrollX = TRUE)
)
```

### Proportion Lab Pending

**Eligibility:** `r afp_prop_lab_pending$metadata$eligibility_note`  
**Threshold:** `r afp_prop_lab_pending$metadata$threshold_rule`

```{r lab-pending-table}
datatable(
  afp_prop_lab_pending$data |>
    dplyr::arrange(ctry) |>
    dplyr::rename(
      Country = ctry,
      Region = whoregion,
      `SG Priority Level` = sg_priority_level,
      `Eligible Samples` = eligible_samples,
      `Pending Samples` = pending_samples,
      `% Pending` = prop_lab_pending,
      Flag = flag
    ),
  rownames=FALSE,
  filter = "top", 
  extensions = "Buttons",
  options = list(
    dom = "Bfrtip", 
    buttons = c("copy", "csv", "excel"),
    pageLength = 15, 
    scrollX = TRUE
    )
)
```

### Timeliness of AFP WPV/VDPV Detection

**Period of focus:** `r afp_wpv_vdpv_timeliness$metadata$recent_period_label`  
**Comparison period:** `r afp_wpv_vdpv_timeliness$metadata$recent_prior_period_label`  
**Earlier period of focus (for context):** `r afp_wpv_vdpv_timeliness$metadata$earlier_period_label`  
**Earlier comparison period (for context):** `r afp_wpv_vdpv_timeliness$metadata$earlier_prior_period_label`  
**Threshold:** `r afp_wpv_vdpv_timeliness$metadata$threshold_rule`

```{r wpv-vdpv-table}
datatable(
  afp_wpv_vdpv_timeliness$data |>
    dplyr::arrange(ctry, current_period) |>
    dplyr::rename(
      Country = ctry,
      Region = whoregion,
      `SG Priority Level` = sg_priority_level,
      Window = window,
      `Current Period` = current_period,
      `Prior Period` = prior_period,
      `Current #` = current_count,
      `Current Median (days)` = current_median_days,
      `Prior #` = prior_count,
      `Prior Median (days)` = prior_median_days,
      `% Change` = perc_change,
      Flag = flag
    ),
  rownames=FALSE,
  filter = "top", 
  extensions = "Buttons",
  options = list(
    dom = "Bfrtip",
    buttons = c("copy", "csv", "excel"),
    pageLength = 15, 
    scrollX = TRUE
    )
)
```

### Timeliness of Detection of Negative AFP Samples

**Period of focus:** `r afp_neg_samples$metadata$current_period_label`  
**Comparison period:** `r afp_neg_samples$metadata$prior_period_label`  
**Threshold:** `r afp_neg_samples$metadata$threshold_rule`

> **Note:** `r afp_neg_samples$metadata$eligibility_note`


```{r neg-samples-table}

 afp_neg_samples$data |>
    dplyr::arrange(country, lubridate::my(month_label)) |>
    dplyr::rename(
      Country = country,
      Region = whoregion,
      `SG Priority Level` = sg_priority_level,
      Month = month_label,
      `Lab Type` = culture.itd.cat,
      `Current #` = current_n,
      `Current Median (days)` = current_median_days,
      `Prior #` = prior_n,
      `Prior Median (days)` = prior_median_days,
      `% Change` = perc_change,
      Flag = flag) |>
  make_sortable_month_col()
```

### Timeliness of Stool Shipment

**Period of focus:** `r afp_timely_stool$metadata$current_period_label`  
**Comparison period:** `r afp_timely_stool$metadata$prior_period_label`  
**Threshold:** `r afp_timely_stool$metadata$threshold_rule`

> **Note:** `r afp_timely_stool$metadata$eligibility_note`


```{r stool-shipment-table}
afp_timely_stool$data |>
    dplyr::arrange(country, lubridate::my(month_label)) |>
    dplyr::rename(
      Country = country,
      Region = whoregion,
      `SG Priority Level` = sg_priority_level,
      Month = month_label,
      `Lab Type` = culture.itd.cat,
      `Current #` = current_n,
      `Current Median (days)` = current_median_days,
      `Prior #` = prior_n,
      `Prior Median (days)` = prior_median_days,
      `% Change` = perc_change,
      Flag = flag
    ) |>
  make_sortable_month_col()

```


### Number of Inadequate Cases

**Definition:**  `r afp_inadequate_cases$metadata$definition`  
**Period of focus:** `r afp_inadequate_cases$metadata$current_period_label`  
**Comparison periods:** `r afp_inadequate_cases$metadata$prior_period_label`  
**Threshold for above or below target:** `r afp_inadequate_cases$metadata$threshold_rule`  

```{r inadequate-cases-table}
afp_inadequate_cases$data |>
    dplyr::arrange(place.admin.0, lubridate::my(month_label)) |>
    dplyr::rename(
      Country = place.admin.0,
      Region = whoregion,
      `SG Priority Level` = sg_priority_level,
      Month = month_label,
      Cases = current_period_counts,
      `3yr Median` = prior_3yr_median,
      `Yrs w/ Data` = prior_yrs_w_data,
      `Percent Change` = perc_change,
      Flag = flag
    ) |>
  make_sortable_month_col()

```


:::

## ES Monitoring Flags

::: panel-tabset


### Proportion of Active ES Sites with Monthly Collections

**Period of focus:** `r es_prop_active_sites_collections$metadata$current_period_label`  
**Threshold:** `r es_prop_active_sites_collections$metadata$threshold_rule`

```{r es-prop-active-collections-table}

  es_prop_active_sites_collections$data |>
    dplyr::arrange(country, lubridate::my(month_label)) |>
    dplyr::rename(
      Country                  = country,
      Region                   = whoregion,
      `SG Priority Level`      = sg_priority_level,
      Month                    = month_label,
      `Active Sites`           = active_sites,
      `Sites w/ ≥1 Collection` = sites_with_1_collection,
      `% With Collection`      = prop_sites_with_1_collection,
      Flag                     = flag
    ) |>
  make_sortable_month_col()

```

### Number of Active ES Sites

**Period of focus:** `r es_active_sites$metadata$current_period_label`  
**Comparison period:** `r es_active_sites$metadata$prior_period_label`  
**Threshold:** `r es_active_sites$metadata$threshold_rule`

```{r es-active-sites-table}

  es_active_sites$data |>
    dplyr::arrange(country, lubridate::my(month_label)) |>
    dplyr::rename(
      Country = country,
      Region = whoregion,
      `SG Priority Level` = sg_priority_level,
      Month = month_label,
      `Current Sites` = current_active_sites,
      `Prior Sites` = prior_active_sites,
      `% Change` = perc_change,
      Flag = flag
    ) |>
  make_sortable_month_col()

```

### Timeliness of ES Shipment

**Period of focus:** `r es_timely_shipment$metadata$current_period_label`  
**Comparison period:** `r es_timely_shipment$metadata$prior_period_label`  
**Threshold:** `r es_timely_shipment$metadata$threshold_rule`

```{r es-shipment-table}
  es_timely_shipment$data |>
    dplyr::arrange(country, lubridate::my(month_label)) |>
    dplyr::rename(
      Country = country,
      Region = whoregion,
      `SG Priority Level` = sg_priority_level,
      Month = month_label,
      `Lab Type` = es.lab.type,
      `Current #` = current_count,
      `Current Median (days)` = current_median_days,
      `Prior #` = prior_count,
      `Prior Median (days)` = prior_median_days,
      `% Change` = perc_change,
      Flag = flag
    ) |>
  make_sortable_month_col()
  
```

### Timeliness of ES WPV/VDPV Notification

**Period of focus:** `r es_wpv_vdpv_timeliness$metadata$recent_period_label`  
**Comparison period:** `r es_wpv_vdpv_timeliness$metadata$recent_prior_period_label`  
**Threshold:** `r es_wpv_vdpv_timeliness$metadata$threshold_rule`

```{r es-wpv-vdpv-table}
datatable(
  es_wpv_vdpv_timeliness$data |>
    dplyr::arrange(country, current_period) |>
    dplyr::rename(
      Country = country,
      Region = whoregion,
      `SG Priority Level` = sg_priority_level,
      Window = window,
      `Current Period` = current_period,
      `Prior Period` = prior_period,
      `Lab Type` = es.lab.type,
      `Current #` = current_count,
      `Current Median (days)` = current_median_days,
      `Prior #` = prior_count,
      `Prior Median (days)` = prior_median_days,
      `% Change` = perc_change,
      Flag = flag
    ),
  rownames=FALSE,
  filter = "top", 
  extensions = "Buttons",
  options = list(
    dom = "Bfrtip", 
    buttons = c("copy", "csv", "excel"),
    pageLength = 15,
    scrollX = TRUE
    )
)
```

:::

## Lab Monitoring Flags

::: panel-tabset

### Timeliness of Virus Isolation

**Period of focus:** `r lab_virus_isolation_timeliness$metadata$current_period_label`  
**Comparison period:** `r lab_virus_isolation_timeliness$metadata$prior_period_label`  
**Threshold:** `r lab_virus_isolation_timeliness$metadata$threshold_rule`

> **Note:** `r lab_virus_isolation_timeliness$metadata$eligibility_note`

```{r virus-isolation-table}

  lab_virus_isolation_timeliness$data |>
    dplyr::arrange(culture.itd.lab, lubridate::my(month_label)) |>
    dplyr::rename(
      `Culture/ITD Lab`      = culture.itd.lab,
      Region = whoregion,
      Month                  = month_label,
      `Current #`            = current_n,
      `Current Median (days)`   = current_median_days,
      `Prior 3yr #`          = prior_3yr_n,
      `Prior 3yr Median (days)` = prior_3yr_median_days,
      `Yrs w/ Data`          = prior_yrs_w_data,
      `% Change`             = perc_change,
      Flag                   = flag
    ) |>
  make_sortable_month_col()

```

### Timeliness of ITD results

**Period of focus:** `r lab_virus_ITD_results_timeliness$metadata$recent_period_label`  
**Comparison period:** `r lab_virus_ITD_results_timeliness$metadata$recent_prior_period_label`  
**Earlier period of focus (for context):** `r lab_virus_ITD_results_timeliness$metadata$earlier_period_label`  
**Earlier comparison period (for context):** `r lab_virus_ITD_results_timeliness$metadata$earlier_prior_period_label`  
**Threshold:** `r lab_virus_ITD_results_timeliness$metadata$threshold_rule`

> **Note:** `r lab_virus_ITD_results_timeliness$metadata$eligibility_note`

```{r ITD-results-table}
datatable(
  lab_virus_ITD_results_timeliness$data |>
    dplyr::arrange(culture.itd.lab, window) |>
    dplyr::rename(
      `Culture/ITD Lab`      = culture.itd.lab,
      Region = whoregion,
      Window = window,
      `Current Period` = current_period,
      `Prior Period` = prior_period,
      `Current #` = current_n,
      `Current Median (days)` = current_median_days,
      `Prior #` = prior_n,
      `Prior Median (days)` = prior_median_days,
      `% Change` = perc_change,
      Flag = flag
    ),
  rownames=FALSE,
  filter = "top",
  extensions = "Buttons",
  options = list(
    dom = "Bfrtip",
    buttons = c("copy", "csv", "excel"),
    pageLength = 15,
    scrollX = TRUE
  )
)
```


### Timeliness of Shipment for Sequencing

**Period of focus:** `r lab_sequencing_shipment_timeliness$metadata$recent_period_label`  
**Comparison period:** `r lab_sequencing_shipment_timeliness$metadata$recent_prior_period_label`  
**Earlier period of focus (for context):** `r lab_sequencing_shipment_timeliness$metadata$earlier_period_label`  
**Earlier comparison period (for context):** `r lab_sequencing_shipment_timeliness$metadata$earlier_prior_period_label`  
**Threshold:** `r lab_sequencing_shipment_timeliness$metadata$threshold_rule`

> **Note:** `r lab_sequencing_shipment_timeliness$metadata$eligibility_note`

```{r sequencing-shipment-table}
datatable(
  lab_sequencing_shipment_timeliness$data |>
    dplyr::arrange(culture.itd.lab, window) |>
    dplyr::rename(
      `Culture/ITD Lab`      = culture.itd.lab,
      Region = whoregion,
      Window                  = window,
      `Current Period` = current_period,
      `Prior Period` = prior_period,
      `Current #` = current_n,
      `Current Median (days)` = current_median_days,
      `Prior #` = prior_n,
      `Prior Median (days)` = prior_median_days,
      `% Change` = perc_change,
      Flag = flag
    ),
  rownames=FALSE,
  filter = "top",
  extensions = "Buttons",
  options = list(
    dom = "Bfrtip",
    buttons = c("copy", "csv", "excel"),
    pageLength = 15,
    scrollX = TRUE
  )
)
```

### Lab Workload

**Period of focus:** `r lab_workload$metadata$current_period_label`  
**Comparison period:** `r lab_workload$metadata$prior_period_label`  
**Threshold:** `r lab_workload$metadata$threshold_rule`

> **Note:** `r lab_workload$metadata$eligibility_note`

```{r lab-workload-table}

  lab_workload$data |>
    dplyr::arrange(culture.itd.lab, lubridate::my(month_label)) |>
    dplyr::rename(
      `Culture/ITD Lab`      = culture.itd.lab,
      Region = whoregion,
      Month                  = month_label,
      `Current #`            = current_n,
      `Prior 3yr Median (days)` = prior_3yr_median,
      `Yrs w/ Data`          = prior_yrs_w_data,
      `% Change`             = perc_change,
      Flag                   = flag
    ) |>
  make_sortable_month_col()

```


### Timeliness of Sequencing Results

**Period of focus:** `r lab_sequencing_timeliness$metadata$recent_period_label`  
**Comparison period:** `r lab_sequencing_timeliness$metadata$recent_prior_period_label`  
**Earlier period of focus (for context):** `r lab_sequencing_timeliness$metadata$earlier_period_label`  
**Earlier comparison period (for context):** `r lab_sequencing_timeliness$metadata$earlier_prior_period_label`  
**Threshold:** `r lab_sequencing_timeliness$metadata$threshold_rule`

> **Note:** `r lab_sequencing_timeliness$metadata$eligibility_note`

```{r lab-sequencing-timeliness-table}
datatable(
  lab_sequencing_timeliness$data |>
    dplyr::arrange(seq.lab, window) |>
    dplyr::rename(
      `Sequencing Lab`      = seq.lab,
      Region = whoregion,
      Window                  = window,
      `Current Period` = current_period,
      `Prior Period` = prior_period,
      `Current #` = current_n,
      `Current Median (days)` = current_median_days,
      `Prior #` = prior_n,
      `Prior Median (days)` = prior_median_days,
      `% Change` = perc_change,
      Flag = flag
    ),
  rownames=FALSE,
  filter = "top",
  extensions = "Buttons",
  options = list(
    dom = "Bfrtip",
    buttons = c("copy", "csv", "excel"),
    pageLength = 15,
    scrollX = TRUE
  )
)
```

:::

:::

```{js enable-js-graphs}
//| echo: false

function remove_all_option() {
    document.getElementById("region-table").getElementsByClassName("selectized")[0].selectize.removeOption("");
    document.getElementById("country-table-m-region").getElementsByClassName("selectized")[0].selectize.removeOption("");
    document.getElementById("country-table-m-flag").getElementsByClassName("selectized")[0].selectize.removeOption("");
    document.getElementById("country-table-q-region").getElementsByClassName("selectized")[0].selectize.removeOption("");
    document.getElementById("country-table-q-flag").getElementsByClassName("selectized")[0].selectize.removeOption("");
    document.getElementById("country-table-n-region").getElementsByClassName("selectized")[0].selectize.removeOption("");
    document.getElementById("country-table-n-flag").getElementsByClassName("selectized")[0].selectize.removeOption("");
    
    document.getElementById("lab-region-table").getElementsByClassName("selectized")[0].selectize.removeOption("");
    document.getElementById("lab-country-table-m-region").getElementsByClassName("selectized")[0].selectize.removeOption("");
    document.getElementById("lab-country-table-m-flag").getElementsByClassName("selectized")[0].selectize.removeOption("");
    document.getElementById("lab-country-table-q-region").getElementsByClassName("selectized")[0].selectize.removeOption("");
    document.getElementById("lab-country-table-q-flag").getElementsByClassName("selectized")[0].selectize.removeOption("");
}

function filter_default() {
    document.getElementById("region-table").getElementsByClassName("selectized")[0].selectize.setValue("AFRO");
    
    document.getElementById("country-table-m-region").getElementsByClassName("selectized")[0].selectize.setValue("AFRO");
    document.getElementById("country-table-m-flag").getElementsByClassName("selectized")[0].selectize.setValue("01. AFP Cases Reported");
    document.getElementById("country-table-q-region").getElementsByClassName("selectized")[0].selectize.setValue("AFRO");
    document.getElementById("country-table-q-flag").getElementsByClassName("selectized")[0].selectize.setValue("02. Proportion 60-Day Follow-Up Completed");
    document.getElementById("country-table-n-region").getElementsByClassName("selectized")[0].selectize.setValue("AFRO");
    document.getElementById("country-table-n-flag").getElementsByClassName("selectized")[0].selectize.setValue("03. Proportion Inadequate Cases Unclassified");
    
        document.getElementById("lab-region-table").getElementsByClassName("selectized")[0].selectize.setValue("AFRO");

    document.getElementById("lab-country-table-m-region").getElementsByClassName("selectized")[0].selectize.setValue("AFRO");
    document.getElementById("lab-country-table-m-flag").getElementsByClassName("selectized")[0].selectize.setValue("13. Timeliness of Virus Isolation");
    document.getElementById("lab-country-table-q-region").getElementsByClassName("selectized")[0].selectize.setValue("AFRO");
    document.getElementById("lab-country-table-q-flag").getElementsByClassName("selectized")[0].selectize.setValue("14. Timeliness of ITD results");
    
    
}
function waitForSelectize(callback, maxTries = 50) {
    let tries = 0;
    const check = setInterval(function() {
        tries++;
        const el = document.getElementById("region-table");
        if (el && el.getElementsByClassName("selectized")[0] && el.getElementsByClassName("selectized")[0].selectize) {
            clearInterval(check);
            callback();
        } else if (tries >= maxTries) {
            clearInterval(check);
            console.warn("Selectize widgets not found after waiting.");
        }
    }, 100);
}

window.onload = function() {
    waitForSelectize(function() {
        remove_all_option();
        filter_default();
    });
}

```
