---
title: "AFP Surveillance Monitoring"
date: today
date-format: "MMMM D, YYYY"
format:
  html:
    embed-resources: true
    page-layout: full
    theme: default
execute:
  echo: false
  message: false
  warning: false
---


```{r setup}
#| echo: false
#| include: false
#| warning: false

library(sirfunctions)
library(DT)
library(tidyverse)
library(plotly)
library(crosstalk)
library(cowplot)

# Set end date for report
end_date <- lubridate::ceiling_date(Sys.Date() %m-% months(1), unit = "month") %m-% days(1)

# Get report data from cache or by sourcing generate_data_for_report.R
use_cache <- FALSE  # set TRUE to skip data reload during QMD development

if (use_cache) {
  load("data_cache/cache.rda")
} else {
  source("generate_data_for_report.R")
}
# Note: build_ and create_ functions are sourced inside generate_data_for_report.R

```

::: panel-tabset

## Overview

End date: **`r format(end_date, "%B %d, %Y")`**

The objective of the monthly SG Monitoring Report is as an early indicator warning for potential disruption in the surveillance system. 
Indicators are presented at the country level across AFP, ES, and Lab. 


## AFP Indicators

::: panel-tabset

### AFP Cases Reported

**Period:** `r afp_cases_reported$metadata$current_period_label`  
**Compared to:** `r afp_cases_reported$metadata$prior_period_label`  
**Threshold:** `r afp_cases_reported$metadata$threshold_rule`

```{r afp-cases-table}
datatable(
  afp_cases_reported$data |>
    dplyr::arrange(place.admin.0, month_label) |>
    dplyr::rename(
      Country = place.admin.0,
      Region = whoregion,
      `SG Priority Level` = sg_priority_level,
      Month = month_label,
      Cases = current_period_counts,
      `3yr Median` = prior_median,
      `Yrs w/ Data` = prior_yrs_w_data,
      `Lower 50%` = lower_50pct,
      `Upper 50%` = upper_50pct,
      Flag = flag
    ),
  filter = "top", 
  extensions = "Buttons",
  options = list(
    dom = "Bfrtip", 
    buttons = c("copy", "csv", "excel"),
    pageLength = 15, 
    scrollX = TRUE
    )
)
```

### Proportion 60-Day Follow-Up

**Recent period:** `r afp_prop_60$metadata$recent_period_label`  
**Earlier period:** `r afp_prop_60$metadata$earlier_period_label`  
**Threshold:** `r afp_prop_60$metadata$threshold_rule`

> **Note:** `r afp_prop_60$metadata$eligibility_note`

```{r prop-60-table}
datatable(
  afp_prop_60$data |>
    dplyr::arrange(ctry, period) |>
    dplyr::rename(
      Country = ctry,
      Region = whoregion,
      `SG Priority Level` = sg_priority_level,
      Period = period,
      `Inadequate Cases` = inad_cases,
      `Got 60d` = got_60day,
      `Prop 60d (%)` = prop_60day,
      Flag = flag
    ),
  filter = "top", 
  extensions = "Buttons",
  options = list(
    dom = "Bfrtip", 
    buttons = c("copy", "csv", "excel"),
    pageLength = 15, 
    scrollX = TRUE
    )
)
```

### Proportion Lab Pending

**Eligibility:** `r afp_prop_lab_pending$metadata$eligibility_note`  
**Threshold:** `r afp_prop_lab_pending$metadata$threshold_rule`

```{r lab-pending-table}
datatable(
  afp_prop_lab_pending$data |>
    dplyr::arrange(ctry) |>
    dplyr::rename(
      Country = ctry,
      Region = whoregion,
      `SG Priority Level` = sg_priority_level,
      `Eligible Samples` = eligible_samples,
      `Pending Samples` = pending_samples,
      `% Pending` = prop_lab_pending,
      Flag = flag
    ),
  filter = "top", 
  extensions = "Buttons",
  options = list(
    dom = "Bfrtip", 
    buttons = c("copy", "csv", "excel"),
    pageLength = 15, 
    scrollX = TRUE
    )
)
```

### Proportion Inadequate Cases Classified

**Eligibility:** `r afp_prop_inad_classified$metadata$eligibility_note`  
**Threshold:** `r afp_prop_inad_classified$metadata$threshold_rule`

```{r prop-classified-table}
datatable(
  afp_prop_inad_classified$data |>
    dplyr::arrange(ctry) |>
    dplyr::rename(
      Country = ctry,
      Region = whoregion,
      `SG Priority Level` = sg_priority_level,
      `Inad Cases` = inad_cases,
      `N Classified` = n_classified,
      `N Unclassified` = n_unclassified,
      `% Unclassified` = prop_unclassified,
      Flag = flag
    ),
  filter = "top", 
  extensions = "Buttons",
  options = list(
    dom = "Bfrtip", 
    buttons = c("copy", "csv", "excel"),
    pageLength = 15, 
    scrollX = TRUE)
)
```

### Timely AFP WPV/VDPV Detection

**Recent period:** `r afp_wpv_vdpv_timeliness$metadata$recent_period_label`  
**Compared to:** `r afp_wpv_vdpv_timeliness$metadata$recent_prior_period_label`  
**Earlier period:** `r afp_wpv_vdpv_timeliness$metadata$earlier_period_label`  
**Compared to:** `r afp_wpv_vdpv_timeliness$metadata$earlier_prior_period_label`  
**Threshold:** `r afp_wpv_vdpv_timeliness$metadata$threshold_rule`

```{r wpv-vdpv-table}
datatable(
  afp_wpv_vdpv_timeliness$data |>
    dplyr::arrange(ctry, current_period) |>
    dplyr::rename(
      Country = ctry,
      Region = whoregion,
      `SG Priority Level` = sg_priority_level,
      Window = window,
      `Current Period` = current_period,
      `Prior Period` = prior_period,
      `Current N` = current_count,
      `Current Median (d)` = current_median_days,
      `Prior N` = prior_count,
      `Prior Median (d)` = prior_median_days,
      `% Change` = perc_change,
      Flag = flag
    ),
  filter = "top", 
  extensions = "Buttons",
  options = list(
    dom = "Bfrtip",
    buttons = c("copy", "csv", "excel"),
    pageLength = 15, 
    scrollX = TRUE
    )
)
```

### Timely Detection of Negative AFP Samples

**Period:** `r afp_neg_samples$metadata$current_period_label`  
**Compared to:** `r afp_neg_samples$metadata$prior_period_label`  
**Threshold:** `r afp_neg_samples$metadata$threshold_rule`

```{r neg-samples-table}
datatable(
  afp_neg_samples$data |>
    dplyr::arrange(country, month_label) |>
    dplyr::rename(
      Country = country,
      Region = whoregion,
      `SG Priority Level` = sg_priority_level,
      Month = month_label,
      `Lab Type` = culture.itd.cat,
      `Current N` = current_n,
      `Current Median (d)` = current_median_days,
      `Prior N` = prior_n,
      `Prior Median (d)` = prior_median_days,
      `% Change` = perc_change,
      Flag = flag
    ),
  filter = "top", 
  extensions = "Buttons",
  options = list(
    dom = "Bfrtip", 
    buttons = c("copy", "csv", "excel"),
    pageLength = 15, 
    scrollX = TRUE
    )
)
```

### Timely Stool Shipment

**Period:** `r afp_timely_stool$metadata$current_period_label`  
**Compared to:** `r afp_timely_stool$metadata$prior_period_label`  
**Threshold:** `r afp_timely_stool$metadata$threshold_rule`

```{r stool-shipment-table}
datatable(
  afp_timely_stool$data |>
    dplyr::arrange(country, month_label) |>
    dplyr::rename(
      Country = country,
      Region = whoregion,
      `SG Priority Level` = sg_priority_level,
      Month = month_label,
      `Lab Type` = culture.itd.cat,
      `Current N` = current_n,
      `Current Median (d)` = current_median_days,
      `Prior N` = prior_n,
      `Prior Median (d)` = prior_median_days,
      `% Change` = perc_change,
      Flag = flag
    ),
  filter = "top", 
  extensions = "Buttons",
  options = list(
    dom = "Bfrtip", 
    buttons = c("copy", "csv", "excel"),
    pageLength = 15, 
    scrollX = TRUE
    )
)
```

:::

## ES Indicators

::: panel-tabset

### Timeliness of ES Shipment

**Period:** `r es_timely_shipment$metadata$current_period_label`  
**Compared to:** `r es_timely_shipment$metadata$prior_period_label`  
**Threshold:** `r es_timely_shipment$metadata$threshold_rule`

```{r es-shipment-table}
datatable(
  es_timely_shipment$data |>
    dplyr::arrange(country, month_label) |>
    dplyr::rename(
      Country = country,
      Region = whoregion,
      `SG Priority Level` = sg_priority_level,
      Month = month_label,
      `Lab Type` = es.lab.type,
      `Current N` = current_count,
      `Current Median (d)` = current_median_days,
      `Prior N` = prior_count,
      `Prior Median (d)` = prior_median_days,
      `% Change` = perc_change,
      Flag = flag
    ),
  filter = "top", 
  extensions = "Buttons",
  options = list(
    dom = "Bfrtip", 
    buttons = c("copy", "csv", "excel"),
    pageLength = 15, 
    scrollX = TRUE
    )
)
```

### Timeliness of ES WPV/VDPV Notification

**Recent period:** `r es_wpv_vdpv_timeliness$metadata$recent_period_label`  
**Compared to:** `r es_wpv_vdpv_timeliness$metadata$recent_prior_period_label`  
**Threshold:** `r es_wpv_vdpv_timeliness$metadata$threshold_rule`

```{r es-wpv-vdpv-table}
datatable(
  es_wpv_vdpv_timeliness$data |>
    dplyr::arrange(country, current_period) |>
    dplyr::rename(
      Country = country,
      Region = whoregion,
      `SG Priority Level` = sg_priority_level,
      Window = window,
      `Current Period` = current_period,
      `Prior Period` = prior_period,
      `Lab Type` = es.lab.type,
      `Current N` = current_count,
      `Current Median (d)` = current_median_days,
      `Prior N` = prior_count,
      `Prior Median (d)` = prior_median_days,
      `% Change` = perc_change,
      Flag = flag
    ),
  filter = "top", 
  extensions = "Buttons",
  options = list(
    dom = "Bfrtip", 
    buttons = c("copy", "csv", "excel"),
    pageLength = 15,
    scrollX = TRUE
    )
)
```


### Proportion of Active ES Sites with Monthly Collections

**Period:** `r es_prop_active_sites_collections$metadata$current_period_label`  
**Threshold:** `r es_prop_active_sites_collections$metadata$threshold_rule`

```{r es-prop-active-collections-table}
datatable(
  es_prop_active_sites_collections$data |>
    dplyr::arrange(country, month_label) |>
    dplyr::rename(
      Country                  = country,
      Region                   = whoregion,
      `SG Priority Level`      = sg_priority_level,
      Month                    = month_label,
      `Active Sites`           = active_sites,
      `Sites w/ ≥1 Collection` = sites_with_1_collection,
      `% With Collection`      = prop_sites_with_1_collection,
      Flag                     = flag
    ),
  filter = "top",
  extensions = "Buttons",
  options = list(
    dom = "Bfrtip",
    buttons = c("copy", "csv", "excel"),
    pageLength = 15,
    scrollX = TRUE
  )
)
```


### Number of Active ES Sites

**Period:** `r es_active_sites$metadata$current_period_label`  
**Compared to:** `r es_active_sites$metadata$prior_period_label`  
**Threshold:** `r es_active_sites$metadata$threshold_rule`

```{r es-active-sites-table}
datatable(
  es_active_sites$data |>
    dplyr::arrange(country, month_label) |>
    dplyr::rename(
      Country = country,
      Region = whoregion,
      `SG Priority Level` = sg_priority_level,
      Month = month_label,
      `Current Sites` = current_active_sites,
      `Prior Sites` = prior_active_sites,
      `% Change` = perc_change,
      Flag = flag
    ),
  filter = "top", 
  extensions = "Buttons",
  options = list(
    dom = "Bfrtip", 
    buttons = c("copy", "csv", "excel"),
    pageLength = 15, 
    scrollX = TRUE)
)
```

:::

## Lab Indicators

*Lab indicators are currently being rebuilt. This section will be updated.*

:::


